Analysing the Performance of Shallow-Water Models with the Roofline Model

### 2016-03-09 docs/papers/roofline_model/gocean_roofline.tex
roofline: add more discussion of SIMD results and correct perf results for 52x52 domain as I actually ran 50x50

The E5-1620 CPU supports two different SIMD instruction sets;
Streaming SIMD Extensions (SSE; flag ``-axSSE4.2'') and Advanced
Vector Extensions (AVX; flag ``-xHost''). SSE supports vector lengths
of 128 bits (two double-precision words) and AVX 256 bits (four
double-precision words). For the problem sizes that largely fit within
L3 or L2 cache, switching on the use of SSE instructions provides a
speed-up of between 1.5 and 1.8 - the factor increasing with problem
size until the WSS of the time-stepping loop begins to fall out of L3
cache. It is therefore likely that the reason this factor is below the
ideal value of 2.0 is the overhead associated with constructing the
loop and handling any ``peel'' or ``remainder'' loops. As the problem
size increases, this overhead (as a fraction of the vectorisable
workload) decreases.

@@

@@

Since the caches of both of the CPUs used are eight-way associative,
we also considered whether we might be experiencing ``cache
thrashing'' whereby different locations in memory are mapped to the
same location in memory.

The inner loop of the kernel where all array accesses are aligned and
the number of distinct streams has been reduced to eight (with a
concommitant reduction in the number of FLOPs):
\begin{verbatim}
do ji = ALIGNMENT+1, M, 1
   rtmp1 = (sshn_u(ji  ,jj ) + hu(ji  ,jj  ))*un(ji  ,jj)
   rtmp2 = (sshn_u(ji-ALIGNMENT,jj ) + hu(ji-ALIGNMENT,jj))* &
           un(ji-ALIGNMENT,jj)
   rtmp3 = (sshn_v(ji ,jj )  + hv(ji  ,jj  ))*vn(ji ,jj)
   ssha(ji,jj) =  (rtmp2 - rtmp1 - rtmp3) / sshn_t(ji,jj)
end do
\end{verbatim}

>>>>

The E5-1620 CPU supports two different SIMD instruction sets;
Streaming SIMD Extensions (SSE; flag ``-axSSE4.2'') and Advanced
Vector Extensions (AVX; flag ``-xHost''). SSE supports vector lengths
of 128 bits (two double-precision words) and AVX 256 bits (four
double-precision words). For the problem sizes that largely fit within
L3 or L2 cache, switching on the use of SSE instructions provides a
speed-up of between 1.5 and 1.8 - the factor increasing with problem
size until the WSS of the time-stepping loop begins to fall out of L3
cache. It is therefore likely that the reason this factor is below the
ideal value of 2.0 is the overhead associated with constructing the
loop and handling any ``peel'' or ``remainder'' loops. As the problem
size increases, this overhead (as a fraction of the vectorisable
workload) decreases. {\bf if this is the case then why is my 128x8 domain
not faster than the 32x32?}

When compiling with AVX instructions rather than SSE, the performance
of the Continuity kernel was typically just below that of the latter
for all of the problem sizes fitting within cache. We attribute this
to the higher overheads associated with managing the longer AVX
vectors.

@@

In what follows, the code modifications were no longer restricted to
those that maintained correctness since we were interested in what
characteristics of the code were limiting its performance.

@@

Since the caches of both of the CPUs used are eight-way associative,
we also considered whether we might be experiencing ``cache
thrashing'' whereby different locations in memory are mapped to the
same location in cache. We therefore constructed a version of the
kernel with the number of distinct streams reduced from twelve to
eight. In order to maintain the same Arithmetic Intensity as the
original kernel we also reduced the number of FLOPs. The inner loop of
the resulting kernel was:
\begin{verbatim}
do ji = ALIGNMENT+1, M, 1
   rtmp1 = (sshn_u(ji  ,jj ) + hu(ji  ,jj  ))*un(ji  ,jj)
   rtmp2 = (sshn_u(ji-ALIGNMENT,jj ) + hu(ji-ALIGNMENT,jj))* &
           un(ji-ALIGNMENT,jj)
   rtmp3 = (sshn_v(ji ,jj )  + hv(ji  ,jj  ))*vn(ji ,jj)
   ssha(ji,jj) =  (rtmp2 - rtmp1 - rtmp3) / sshn_t(ji,jj)
end do
\end{verbatim}

Figure~\ref{FIG_cont_vec} shows that the absolute performance of this
modified kernel (``8 arrays'') is mostly lower than that of the kernel
with twelve streams and certainly never exceeds it. We therefore
conclude that cache-thrashing is not a problem for this kernel.
